---
title: "wordle"
format: html
editor: visual
---

## Wordle Challenge

Intro: packages

```{r}
#| eval: false
#| warning: false

library(tidyverse) 
library(dplyr)
```

Step 1: Create Lists

Create your own custom function called `load_dictionary()` that takes a single argument, “filename”, that can be used to read in either of the two data files your downloaded.

```{r}
#| eval: false
collins <- read.table("collins-scrabble-words-2019.txt", header = TRUE, sep = " ")
str(collins)
google <- read.table("google-10000-english-usa-no-swears.txt", header = TRUE, sep = " ")
str(google)
class(google)

load_dictionary <- function(filename) {
  as.character(filename$words)
}

valid_list <- load_dictionary(collins)
str(valid_list) #279496 words

solution_list <- load_dictionary(google)
str(solution_list) #9884 words
```

Step 2: Ensure all solutions are also valid

```{r}
#| eval: false

?intersection
solution_list <- intersect(solution_list, valid_list)
str(solution_list) #8336 words
```

Step 3: Create function to pick a solution word

```{r}
pick_solution <- function(word_list) {
  word_list <- word_list[nchar(word_list) == 5] 
  solution <- sample(word_list, 1) 
  solution_letters <- strsplit(solution, "") [[1]]
  return(solution_letters)
}

solution <- pick_solution(solution_list)
solution

#in plain language: for each string in vector word_list, count how many letters are in the word, if == 5 keep, if not, drop
#alt: filter out all words that != 5 letters

#it works! returns a vector of 5 letters when run like -> pick_solution(solution_list)
```

Step 4: Functions needed to play

What does the function need to do?

\[1\] tell the player the rules of the game, \[2\] display what letters the player has not yet guessed, \[3\] Prompt the player for a guess, \[4\] read in their guess, and check that their guess is valid, \[5\] Compare the guess to the solution word and generate the necessary feedback, e.g., `*` for *in the word and in the correct position*, `+` for *in the word but in the wrong position*, and `-` for *not in the word*., \[6\] update the list of letters not yet guessed, \[7\] print success if they get it, \[8\] print failure after 6 wrong guesses

```{r}
#making evaluate guess function that will check if any letters are in the right place
evaluate_guess <- function(guess, solution){
  outcome <- c("-", "-", "-", "-", "-")
  guess <- toupper(guess)
  guess_letters <- strsplit(guess, "")[[1]]
  for(i in 1:5){
  if (guess_letters[i] == solution[1]) {outcome[i] <- "+"}
  if (guess_letters[i] == solution[2]) {outcome[i] <- "+"}
  if (guess_letters[i] == solution[3]) {outcome[i] <- "+"}
  if (guess_letters[i] == solution[4]) {outcome[i] <- "+"}
  if (guess_letters[i] == solution[5]) {outcome[i] <- "+"}
  }
  for(i in 1:5){
    if (guess_letters[i] == solution[i]) {outcome[i] <- "*"}
  }
  return(outcome)
}
  }
```

```{r}
#play wordle needs to...

#define alphabet
remaining_letters <- LETTERS

#establish a solution (don't tell!)
solution <- pick_solution(solution_list)

#explain the rules of the game 
print("Input a 5 letter word.")

#get a guess from the user
print(remaining_letters)
guess <- readline(prompt = "Guess: ")

#evaluate guess
evaluate_guess(guess, solution)
remaining_letters <- remaining_letters[!(remaining_letters %in% guess_letters)]
#this will return the outcome vector

#check if the last guess was a winner
if(outcome != c("*", "*", "*", "*", "*")) {print("Try Again!")}
if(outcome == c("*", "*", "*", "*", "*")) {print("You Win!")} #and end the loop

```

```{r}
play_wordle <- function() {
  remaining_letters <- LETTERS
  solution <- pick_solution(solution_list)
  count = 0
  while(outcome != c("*", "*", "*", "*", "*") || count < 6)
  {
    print(remaining_letters)
    guess <- readline(prompt = "Guess: ")
    evaluate_guess(guess, solution)
    remaining_letters <- remaining_letters[!(remaining_letters %in% guess_letters)]
    if(outcome != c("*", "*", "*", "*", "*")) {print("Try Again!")}
  }
  count = count + 1
  if(outcome == c("*", "*", "*", "*", "*")) {print("You Win!")}
  if(count == 6) {print(paste0("You are out of guesses! The answer was:", solution))}
}


```

