---
title: "wordle"
format: html
editor: visual
---

## Wordle Challenge

Intro: packages

```{r}
#| eval: false
#| warning: false

library(tidyverse) 
library(dplyr)
```

Step 1: Create Lists

Create your own custom function called `load_dictionary()` that takes a single argument, “filename”, that can be used to read in either of the two data files your downloaded.

```{r}
#| eval: false
collins <- read.table("collins-scrabble-words-2019.txt", header = TRUE, sep = " ")
str(collins)
google <- read.table("google-10000-english-usa-no-swears.txt", header = TRUE, sep = " ")
str(google)
class(google)

load_dictionary <- function(filename) {
  as.character(filename$words)
}

valid_list <- load_dictionary(collins)
str(valid_list) #279496 words

solution_list <- load_dictionary(google)
str(solution_list) #9884 words
```

Step 2: Ensure all solutions are also valid

```{r}
#| eval: false

?intersection
solution_list <- intersect(solution_list, valid_list)
str(solution_list) #8336 words
```

Step 3: Create function to pick a solution word

```{r}
#| eval: false

pick_solution <- function(word_list) {
  word_list <- word_list[nchar(word_list) == 5] 
  solution <- sample(word_list, 1) 
  solution_letters <- strsplit(solution, "") [[1]]
  return(solution_letters)
}

solution <- pick_solution(solution_list)
solution

#it works! returns a vector of 5 letters when run like -> pick_solution(solution_list)
```

Step 4: Functions needed to play

What does the function need to do?

\[1\] tell the player the rules of the game, \[2\] display what letters the player has not yet guessed, \[3\] Prompt the player for a guess, \[4\] read in their guess, and check that their guess is valid, \[5\] Compare the guess to the solution word and generate the necessary feedback, e.g., `*` for *in the word and in the correct position*, `+` for *in the word but in the wrong position*, and `-` for *not in the word*., \[6\] update the list of letters not yet guessed, \[7\] print success if they get it, \[8\] print failure after 6 wrong guesses

```{r}
#making evaluate guess function that will check if any letters are in the right place
evaluate_guess <- function(guess, solution){

guess_letters <- strsplit(guess, "")[[1]]
solution_counts <- table(solution)
n <- length(solution)
outcome <- rep("-", n)

    for(i in 1:n)
      {
      if (guess_letters[i] == solution[i]) 
        {
        outcome[i] <- "*"
        solution_counts[guess_letters[i]] <- solution_counts[guess_letters[i]] - 1
      }
    }

for(i in 1:n)
      if (outcome[i] == "-") {
        letter <- guess_letters[i]
        
        if(letter %in% names(solution_counts) && solution_counts[letter] > 0) {
          outcome[i] <- "+"
          solution_counts[letter] <- solution_counts[letter] - 1
      }
      }
  return(outcome)
}
```

```{r}
#| eval: false

play_wordle <- function(solution_list, valid_list, num_guesses) {

  #pick a solution, split it into a vector of letters
  solution <- pick_solution(solution_list)

  #set variables to hold things
  remaining_letters <- LETTERS
  x <- 0
  n <- length(solution)

 #intro 
  cat("Welcome to Wordle! You will have 6 tries to guess the secret word. After each guess, you will get a series of symbols that tell you how well you are doing. * means the letter was correct, and in the correct spot. + means the letter if in the word, but not in the spot you guessed. - means the letter is not in the word.\n")

  #loop
  while(any(outcome != "*") && x < num_guesses) {
    cat("Remaining letters :\n", remaining_letters, "\n")
    
    guess <- toupper(readline(prompt = "Guess a 5-letter word: "))
    guess_letters <- strsplit(guess, "")[[1]]
    
    ##check that guess is the right length
      if(!(nchar(guess) == 5)) {
      cat("Your guess must contain ", n, "letters. Try again:\n")
        next
      }
    
    ##check that guess is valid
    if (!(guess %in% valid_list)) {
    cat("This word is invalid. Try again: \n")
      next
      }

# evaluate a valid guess
    outcome <- evaluate_guess(guess, solution)
    print(outcome)
    remaining_letters <- remaining_letters[!remaining_letters %in% guess_letters]
    x <- x + 1
  }
  
  #end of game
  if(all(outcome == "*")) {
    print("You Win!")
  } else {
    cat("You are out of guesses! The answer was:", solution)
  }
}
```

```{r}
#| eval: false

#run this code to play (with 6 tries)!
#I am not sure how you want to be able to test it, or if there is a way to initialize more naturally with quarto. Sorry if this is harder than it needs to be to play!
play_wordle(solution_list, valid_list, 6)

```

